Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}

### 2017-08-06 ident_manu.tex
new table with flask calibration, not final

\title{Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Version: Thu Aug 3 12:09:09 2017 -0500, \href{https://github.com/fawda123/identifiability/commit/0895cb42311eb3eb06b254be874f61997b2288d5}{0895cb42311eb3eb06b254be874f61997b2288d5}\newline Acronyms: \acfi{chla}; \acfi{cgem}; \acfi{do}; \acfi{dom}; \acfi{gom}; \acfi{lcs}; \acfi{marb}; \acfi{pom}; \acfi{rmse}; \acfi{zerod}}

@@

Calibration of parameter values for each of the seven selection heuristics (\cref{tab:heurist1,tab:heurist2,tab:heurist3})  produced model output that varied in goodness of fit and computational effort required to identify optimal values (\cref{tab:calib}).  As expected, the number of required iterations of the optimization algorithm to minimize \ac{rmse} of observed \ac{do} and model output varied in proportion to the number of parameters in a subset.  For example, only six iterations were required to optimize the single parameter selected for the temperature category, whereas 1530 iterations were required for the eight parameters selected independent of category.  No association was observed between collinearity ($\gamma$) and number of iterations or calibration.  Of the seven parameter subsets, selection independent of parameter category and equally within each category produced the largest reduction in \ac{rmse} values from the default output (starting \ac{rmse} $72.14$ mmol O$_2$ m$^{-3}$, reduction to $66.77$, $67.44$, respectively), whereas the optimal values for parameters in the Optics category had the smallest reduction ($72.12$).  Lower \ac{rmse} was associated with higher average sensitivity of a parameter subset.  Parameters selected only in the optics category had the highest \ac{rmse} and lowest average L1 ($2.81\times 10^{-4}$, three parameters), whereas parameters selected independent of category had the lowest \ac{rmse} and highest average L1 ($0.03$, eight parameters, \cref{tab:calib}).

@@

% calibration table, do
%latex.default(totab[, -1], file = "", rowlabel = "Selection heuristic",     caption = cap_val, caption.loc = "top", size = "scriptsize",     label = "tab:calib", rowname = totab$heurist, rgroup = c("1",         "2", "3"), insert.bottom = foot.val, n.rgroup = c(5,         1, 1), colheads = colnms)%
\begin{table}[!tbp]
{\scriptsize
\caption{Results of model calibration to observed \ac{do} for each of seven parameter subsets selected by sensitivity within each parameter category (first), independent of category (second), and equally within each category (third). The collinearity values ($\gamma$) and parameters of each subset are from \cref{tab:heurist1,tab:heurist2,tab:heurist3}.  Mean L1 is the average of sensitivity values (\cref{l1}) for the subset.  $RMSE$ shows the final error comparing observed and model output after the specified number of iterations for optimization.\label{tab:calib}} 
\begin{center}
\begin{tabular}{llllll}
\hline\hline
\multicolumn{1}{l}{Selection heuristic}&\multicolumn{1}{c}{$\gamma$}&\multicolumn{1}{c}{$n$ parameters}&\multicolumn{1}{c}{mean L1}&\multicolumn{1}{c}{Iterations}&\multicolumn{1}{c}{$RMSE$*}\tabularnewline
\hline
{\bfseries 1}&&&&&\tabularnewline
~~Optics&$3.87$&$3$&$2.81\times 10^{-4}$&$49$&$72.12$\tabularnewline
~~Organic Matter&$8.61$&$5$&$3.18\times 10^{-3}$&$154$&$68.7$\tabularnewline
~~Phytoplankton&$3.97$&$7$&$0.02$&$1725$&$68.3$\tabularnewline
~~Temperature&$1$&$1$&$0.02$&$6$&$71.76$\tabularnewline
~~Zooplankton&$5.2$&$4$&$0.03$&$279$&$71.22$\tabularnewline
\hline
{\bfseries 2}&&&&&\tabularnewline
~~Independent of category&$6.81$&$8$&$0.03$&$1530$&$66.77$\tabularnewline
\hline
{\bfseries 3}&&&&&\tabularnewline
~~Equally within category&$5.22$&$7$&$0.03$&$1005$&$67.44$\tabularnewline
\hline
\end{tabular}\end{center}}
\footnotesize *Starting \ac{rmse} $= 72.140$\end{table}

>>>>

\title{Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Version: Fri Aug 4 17:59:08 2017 -0500, \href{https://github.com/fawda123/identifiability/commit/84e94b5eb0aca29df79344221e5651a5af428017}{84e94b5eb0aca29df79344221e5651a5af428017}\newline Acronyms: \acfi{chla}; \acfi{cgem}; \acfi{do}; \acfi{dom}; \acfi{gom}; \acfi{lcs}; \acfi{marb}; \acfi{pom}; \acfi{rmse}; \acfi{zerod}}

@@

Calibration of parameter values for each of the seven selection heuristics (\cref{tab:heurist1,tab:heurist2,tab:heurist3})  produced model output that varied in goodness of fit and computational effort required to identify optimal values (\cref{tab:calib}).  As expected, the number of required iterations of the optimization algorithm to minimize \ac{rmse} of observed \ac{do} and model output varied in proportion to the number of parameters in a subset.  For example, only six iterations were required to optimize the single parameter selected for the temperature category, whereas 136 iterations were required for the eight parameters selected independent of category.  No association was observed between collinearity ($\gamma$) and number of iterations or calibration.  Of the seven parameter subsets, selection independent of parameter category and equally within each category produced the largest reduction in \ac{rmse} values from the default output (starting \ac{rmse} $33.47$ mmol O$_2$ m$^{-3}$, reduction to $26.56$, $28.36$, respectively), whereas the optimal values for parameters in the Optics category had the smallest reduction ($33.3$).  Lower \ac{rmse} was associated with higher average sensitivity of a parameter subset.  Parameters selected only in the optics category had the highest \ac{rmse} and lowest average L1 ($2.81\times 10^{-4}$, three parameters), whereas parameters selected independent of category had the lowest \ac{rmse} and highest average L1 ($0.03$, eight parameters, \cref{tab:calib}).

@@

% calibration table, do
%latex.default(totab[, -1], file = "", rowlabel = "Selection heuristic",     caption = cap_val, caption.loc = "top", size = "scriptsize",     label = "tab:calib", rowname = totab$heurist, rgroup = c("1",         "2", "3"), insert.bottom = foot.val, n.rgroup = c(5,         1, 1), colheads = colnms)%
\begin{table}[!tbp]
{\scriptsize
\caption{Results of model calibration to observed \ac{do} for each of seven parameter subsets selected by sensitivity within each parameter category (first), independent of category (second), and equally within each category (third). The collinearity values ($\gamma$) and parameters of each subset are from \cref{tab:heurist1,tab:heurist2,tab:heurist3}.  Mean L1 is the average of sensitivity values (\cref{l1}) for the subset.  $RMSE$ shows the final error comparing observed and model output after the specified number of iterations for optimization.\label{tab:calib}} 
\begin{center}
\begin{tabular}{llllll}
\hline\hline
\multicolumn{1}{l}{Selection heuristic}&\multicolumn{1}{c}{$\gamma$}&\multicolumn{1}{c}{$n$ parameters}&\multicolumn{1}{c}{mean L1}&\multicolumn{1}{c}{Iterations}&\multicolumn{1}{c}{$RMSE$*}\tabularnewline
\hline
{\bfseries 1}&&&&&\tabularnewline
~~Optics&$3.87$&$3$&$2.81\times 10^{-4}$&$49$&$33.3$\tabularnewline
~~Organic Matter&$8.61$&$5$&$3.18\times 10^{-3}$&$44$&$31.48$\tabularnewline
~~Phytoplankton&$3.97$&$7$&$0.02$&$300$&$25.84$\tabularnewline
~~Temperature&$1$&$1$&$0.02$&$6$&$31.23$\tabularnewline
~~Zooplankton&$5.2$&$4$&$0.03$&$72$&$32.99$\tabularnewline
\hline
{\bfseries 2}&&&&&\tabularnewline
~~Independent of category&$6.81$&$8$&$0.03$&$136$&$26.56$\tabularnewline
\hline
{\bfseries 3}&&&&&\tabularnewline
~~Equally within category&$5.22$&$7$&$0.03$&$45$&$28.36$\tabularnewline
\hline
\end{tabular}\end{center}}
\footnotesize *Starting \ac{rmse} $= 72.140$\end{table}
